Structured Versus Decorated Cospans

### 2021-06-22 cospans_new.tex
corrected a lot of problems Kenny found in his email

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$  whenever $A$ is a cocartesian.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   These results build on earlier work of Fong \cite{Fong}, and we recommend his thesis for intuitive explanations of the fundamental ideas \cite{FongThesis}.

@@

\[  \one\cong\one \times \one \xrightarrow{s \, \times \, t} F(n) \times F(m) \xrightarrow{\phi_{n,m}} F(m+n) \xrightarrow{F(\psi)} F(m +_{b}n). \]

@@

\begin{thm}\label{thm:decorated_cospans_2}
Let $(\A,+)$ be a cocartesian monoidal category and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ be a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a \to m \leftarrow b, s \in F(m))$ and $(c \to n \leftarrow d, t \in F(n))$ is
\[
\begin{tikzcd}
& m+n & \\
a+c\ar[ur,"i_1+i_2"] && b+d,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad s \otimes t:=  \phi_{m,n}(s,t) \in F(m + n) 
\]
\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(s) \to s' \textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(t) \to t'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m,n}(s,t))\to \phi_{m',n'}(s',t')\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"s\,\times \, t"]\ar[dr,"s' \,\times \, t' "']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in \C$ with $\Phi(c) = a + b$.
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a Beck--Chevalley symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

@@

Note that when $\C$ is the arrow category of $\A$ and $\Phi$ maps any arrow in $\C$ to its
domain, an object of $\Fr(\Phi)$ simply amounts to a cospan in $\A$.  In this case $\Fr(\Phi)$ is the double category of cospans in $\A$.    We shall use variations on this idea to construct double categories of decorated and structured cospans, and to prove that under certain conditions these double categories are equivalent.    We begin with decorated cospans, proving Theorem \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying Shulman's result to a particular monoidal bifibration built from a lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$.

@@

Moreover, each fiber $\X_a=F(a)$ obtains a monoidal structure via 
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a)
\end{equation}
where $\nabla$ is the fold map, which is symmetric when $F$ is, again via the components of $u_{a,a}$. Also, each reindexing functor $f_!=F(f)$ obtains a strong monoidal structure with
\begin{equation}\label{eq:strongreindexing}
\begin{tikzcd}[row sep=.35in]
 F(a)\times F(a)\ar[r,"\phi_{a,a}"]\ar[d,"F(f)\times F(f)"']\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{f,f}}{\cong}}"] & 
 F(a+a)\ar[r,"F(\nabla)"]\ar[d,"F(f+f)"description]\ar[dr,phantom,"\cong"] & F(a)\ar[d,"Ff"] \\
 F(b)\times F(b)\ar[r,"\phi_{b,b}"'] & F(b+b)\ar[r,"F(\nabla)"'] & F(b)
\end{tikzcd}
\end{equation}
\end{proof}

@@

Next, one can check that $\Phi$ is a fibration.  Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a' \xrightarrow{i'} b', s') \in \A/U$ over $a'$, a cartesian lifting of $f$ to this object is given by 
\[ (f,1,1) \maps (a \xrightarrow{i' f} b' , s')\; \to \; (a' \xrightarrow{i'} b', s'). \]
Denoting the cartesian lifting of $f$ to an object $z \in \A/U$ by $\mathrm{Cart}(f,z)$, the tensor product of $\A/U$ strictly preserves such cartesian liftings in the sense that $\mathrm{Cart}(f_1,z_1) \otimes \mathrm{Cart}(f_2,z_2)=\mathrm{Cart}(f_1 \otimes f_2, z_1 \otimes z_2)$.

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, s) \in \A/U$ over $a$, we claim that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, s) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(s)). \]
where $g, i'$ and $b +_a a'$ arise from this pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$.};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
Suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, s) \; \to \; (a'' \xrightarrow{i''} b'', s'')   \]
in $\A/U$ and a morphism $k \maps a \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k f$.   To show that $(f,g,1)$ is a cocartesian lifting of $f$, we need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(s)) \; \to \; (a'' \xrightarrow{i''} b'', s'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this we choose $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[below left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition obeyed by morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

It follows that $\Phi$ is a bifibration.  Denoting the cocartesian lifting of a morphism $f$ to an object $z \in \A/U$ by $\mathrm{Cocart}(f,z)$, we have that $\mathrm{Cocart}(f_1,z_1) \otimes \mathrm{Cocart}(f_2,z_2) \cong \mathrm{Cocart}(f_1 \otimes f_2,z_1 \otimes z_2)$ by the universal map between two colimits of the same diagram.    Thus, $\Phi$ is a monoidal bifibration.

Lastly, we need to check that $\Phi$ satisfies the Beck--Chevalley condition.  Merely from the fact that $\Phi$ is a bifibration, for any commutative square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\label{bc-square}
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
there is a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$.};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
defined in \eqref{theta}.   We need to prove that when the square in $\A$ is 
a pushout, $\theta$ is a natural isomorphism.

@@

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\|H \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

@@

The comma category $L/\X$ has finite colimits because $A$ and $X$ have finite colimits and $L$ 
preserves them.   It thus becomes symmetric monoidal with the chosen finite coproducts providing the monoidal structure.   The tensor product of two objects $(a,L(a) \xrightarrow{i} x)$ and $(a,L(a) \xrightarrow{i'} x')$ is given by 
\[  (a,L(a) \xrightarrow{i} x) + (a',L(a') \xrightarrow{i'} x')=(a+a',x+x',L(a+a')  \xrightarrow{\sim} L(a)+L(a') \xrightarrow{i+i'} x+x'). \]
The monoidal unit is ${(0_\A,0_\X,{L(0_\A) \to 0_\X})}$.  It is clear that $\Psi \maps (L/\X, +) \to (\A,+)$ is a strict monoidal functor.

>>>>

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ whenever $A$ has finite colimits.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   These results build on earlier work of Fong \cite{Fong}, and we recommend his thesis for intuitive explanations of the fundamental ideas \cite{FongThesis}.

@@

\[  \one\cong\one \times \one \xrightarrow{s \, \times \, t} F(m) \times F(n) \xrightarrow{\phi_{m,n}} F(m+n) \xrightarrow{F(\psi)} F(m +_{b}n). \]

@@

\begin{thm}\label{thm:decorated_cospans_2}
Let $\A$ be a category with finite colimits and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ be a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ of \cref{thm:decorated_cospans_1} is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a \xrightarrow{i_1} m \xleftarrow{o_1} b, s \in F(m))$ and $(c \xrightarrow{i_2} n \xleftarrow{o_2} d, t \in F(n))$ is
\[
\begin{tikzcd}
& m+n & \\
a+c\ar[ur,"i_1+i_2"] && b+d,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad s \otimes t:=  \phi_{m,n}(s,t) \in F(m + n) 
\]
\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(s) \to s' \textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(t) \to t'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m,n}(s,t))\to \phi_{m',n'}(s',t')\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"s\,\times \, t"]\ar[dr,"s' \,\times \, t' "']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in \C$ with $\Phi(c) = a + b$,
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a Beck--Chevalley symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

@@

Note that when $\C$ is the arrow category of $\A$ and $\Phi$ maps any arrow in $\C$ to its
domain, a horizontal 1-cell $\Fr(\Phi)$ simply amounts to a cospan in $\A$.  In this case $\Fr(\Phi)$ is the double category of cospans in $\A$.    We shall use variations on this idea to construct double categories of decorated and structured cospans, and to prove that under certain conditions these double categories are equivalent.    We begin with decorated cospans, proving Theorem \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying Shulman's result to a particular monoidal bifibration built from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$.

@@

Moreover, each fiber $\X_a=F(a)$ obtains a monoidal structure via 
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a)
\end{equation}
where $\nabla$ is the fold map, which is symmetric when $F$ is, again via the components of $u_{a,a}$. Also, each reindexing functor $f_!=F(f)$ obtains a strong monoidal structure with
\begin{equation}\label{eq:strongreindexing}
\begin{tikzcd}[row sep=.35in]
 F(a)\times F(a)\ar[r,"\phi_{a,a}"]\ar[d,"F(f)\times F(f)"']\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{f,f}}{\cong}}"] & 
 F(a+a)\ar[r,"F(\nabla)"]\ar[d,"F(f+f)"description]\ar[dr,phantom,"\cong"] & F(a)\ar[d,"F(f)"] \\
 F(b)\times F(b)\ar[r,"\phi_{b,b}"'] & F(b+b)\ar[r,"F(\nabla)"'] & F(b)
\end{tikzcd}
\end{equation}
\end{proof}

@@

Next, one can check that $\Phi$ is a fibration.  Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a' \xrightarrow{i'} b', s') \in \A/U$ over $a'$, a cartesian lifting of $f$ to this object is given by 
\[ (f,1,1) \maps (a \xrightarrow{i' f} b' , s')\; \to \; (a' \xrightarrow{i'} b', s'). \]
Denoting the cartesian lifting of $f$ to an object $z \in \A/U$ by $\mathrm{Cart}(f,z)$, the tensor product of $\A/U$ strictly preserves such cartesian liftings in the sense that $\mathrm{Cart}(f_1,z_1) \otimes \mathrm{Cart}(f_2,z_2)=\mathrm{Cart}(f_1 + f_2, z_1 \otimes z_2)$.

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, s) \in \A/U$ over $a$, we claim that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, s) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(s)). \]
where $g, i'$ and $b +_a a'$ arise from this pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$.};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
To show that $(f,g,1)$ is a cocartesian lifting of $f$, suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, s) \; \to \; (a'' \xrightarrow{i''} b'', s'')   \]
in $\A/U$ and a morphism $k \maps a' \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k f$.   We need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(s)) \; \to \; (a'' \xrightarrow{i''} b'', s'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this we choose $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[below left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition obeyed by morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

It follows that $\Phi$ is a bifibration.  Denoting the cocartesian lifting of a morphism $f$ to an object $z \in \A/U$ by $\mathrm{Cocart}(f,z)$, we have that $\mathrm{Cocart}(f_1,z_1) \otimes \mathrm{Cocart}(f_2,z_2) \cong \mathrm{Cocart}(f_1 + f_2,z_1 \otimes z_2)$ by the universal map between two colimits of the same diagram.    Thus, $\Phi$ is a monoidal bifibration.

Lastly, we need to check that $\Phi$ satisfies the Beck--Chevalley condition.  Merely from the fact that $\Phi$ is a bifibration, for any commutative square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\label{bc-square}
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
there is a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
defined in \eqref{theta}.   We need to prove that when the square in $\A$ is 
a pushout, $\theta$ is a natural isomorphism.

@@

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a monoidal natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\lH \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

@@

The comma category $L/\X$ has finite colimits because $A$ and $X$ have finite colimits and $L$ 
preserves them.   It thus becomes symmetric monoidal with the chosen finite coproducts providing the monoidal structure.   The tensor product of two objects $(a,L(a) \xrightarrow{i} x)$ and $(a,L(a) \xrightarrow{i'} x')$ is given by 
\[  (a,L(a) \xrightarrow{i} x) + (a',L(a') \xrightarrow{i'} x')=(a+a',L(a+a')  \xrightarrow{\sim} L(a)+L(a') \xrightarrow{i+i'} x+x'). \]
The monoidal unit is ${(0_\A,0_\X,{L(0_\A) \xrightarrow{!} 0_\X})}$.  It is clear that $\Psi \maps (L/\X, +) \to (\A,+)$ is a strict monoidal functor.

The functor $\Psi$ is both a fibration and an opfibration. Suppose we are given a morphism $f \maps a \to b$ in $\A$.  For any object $(b,L(b) \xrightarrow{i} x) \in L/\X$ over $b$, a cartesian lifting of $f$ is given by
\[ (f,1_x) \maps (a,L(a) \xrightarrow{i L(f)} x) \; \to \; (b,L(b) \xrightarrow{i} x) . \]
For any object $(a,L(a) \xrightarrow{i} x)$ in $L/\X$ over $a$, a cocartesian lifting of $f$ is given by
\[ (f,\kappa) \maps (a,L(a) \xrightarrow{i} x) \to (b,L(b) \xrightarrow{\theta} x+_{L(a)} L(b)) \]
where $\kappa$ and $\theta$ arise from the pushout in this diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (2.25,0) {$L(c)$};
\node (C) at (1.5,1) {$L(b)$};
\node (D) at (0,-2) {$x$};
\node (E) at (2.25,-2) {$y$};
\node (F) at (1.5,-1) {$x+_{L(a)} L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge [dashed] node [above] {$\kappa$} (F)
(F) edge [dashed] node [left] {$\exists ! \psi$} (E)
(D) edge node [above] {$g'$} (E)
(A) edge node [left] {$i$} (D)
(C) edge [dashed] node [left,near end] {$\theta$} (F)
(B) edge node {$$} (E)
(A) edge node[above]{$L(f')$} (B)
(C) edge node [right] {$L(p)$} (B)
(A) edge node[left]{$L(f)$} (C);
\end{tikzpicture}
\]
This diagram also gives the proof that $(f,\kappa)$ is a cocartesian lifting of $f$.
Both cartesian and cocartesian liftings are clearly preserved by the tensor product of $L/\X$, so
$\Psi$ is a monoidal bifibration.
